---
title: "internship_report"
author: "Junhao Wang"
date: "26/05/2022"
output:
  bookdown::html_document2: default
---

# Amber Community Exploratory Data Analysis Program

### Abstract

This program is offered by Leanne Hyndman, the leader of Amber Community, and aims to find useful information through the available data provided by the team or other contributors on the internet and provide data-driven support.

To achieve this aim, I cleaned the data from Amber Community, joined it with other data, did some exploration analysis, made the time series analysis, and compared the population and referrals spatial distribution.

The data contains the following variables: 

* Date received, the date Amber Community received the referrals
* Date entered, the date the client entered
* Client Type,
* Referred by, the type of the person or institution that made the referral
* Postcode,
* DHA Region, which looks a bit messy and not format is not unified

There are mainly two findings: one is that the number of referrals during the time shows a weak seasonality weekly and monthly, which may be due to the regular schedule of staff. The other is that the institution seems to have different levels of “attractiveness” to different regions: the people who live in regional areas seem to be more likely to come to the institution.

/The data is from three sources: Amber Community for the essential data, the r package called “adsmapsdata” for the map data, and the government's website for demographics(address: [https://www.coronavirus.vic.gov.au/victorian-coronavirus-covid-19-data](https://www.coronavirus.vic.gov.au/victorian-coronavirus-covid-19-data)).



### Background and Motivation

Amber Community, formerly Road Trauma Support Services Victoria, is a not-for-profit organization contributing to the safety and well-being of road users. To provide data-driven support. 

They provide counselling and support to people affected by road trauma and address the attitudes and behaviours of road users through education. Also, They deliver a range of education programs addressing the behaviours and attitudes of drivers to reduce the incidence of crashes, injuries and fatalities, and the associated trauma and grief.

Now, they are interested in a few questions and wish that we can provide some data-driven support so that they can have more insight into their meaningful work and contribute more to road users’ mental and physical health.

### Objectives and Significance

* Summarise the data and note any changes over the period:
+ referred by
+ number of days between date received and date entered
+ client type
+ number of referrals per day/week/month

* Hopefully, see any seasonality

* Figure out how do the location demographics compare to the population of VIC?

A tricky question here is to compare the demographics, which I considered to use the formula:

 $$\frac{a_1}{a_2}-\frac{b_1}{b_2}$$


### Data Exploration

The data looks tidy but a bit messy with a lot of missing values. With the help of my mentor Rob Hyndman, I removed useless variables and cleaned the variable names and created the table referrals_clean1.

Then I checked the data type, name and value in each variable using the function glimpse():

After checking that, I renamed the variable and combined the categorical variables' values that are identical at a certain level and then created a new table called referrals_clean2. 

Next, I created a new variable "day_of_week" to check whether more referrals are received on Monday, and the result shows that Monday does come with more referrals, which indicates that there are possibilities of seasonality. According to the table above, there are more referrals on Monday than on other days of the week. This might be because usually some referrals on weekends are postponed to the following day, which is Monday, and this can also explain why Tuesday has the second-largest amount of referrals, and the rest of the weekdays(Wednesday, Thursday and Friday) have about the same amount, which is still significantly larger(around 40%) than the weekend.

I had a look at the summary of the client_type variable:

From the table above we can see that the major source of clients is from drivers, witnesses and bereaved people.

I also had a look at the summary of the referred_by variable:

And it is clear that VPeR is the major source of referrals.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = TRUE)
```

```{r}
# Load tidyverse packages
library(tidyverse)
library(readr)
library(janitor)
library(stringr)
library(lubridate)
library(fpp3)
library(absmapsdata)
library(sf)
library(leaflet)
library(RColorBrewer)
library(plotly)
library(ggiraph)
library(knitr)
```

```{r}
# Read files
referrals <- read_csv(file="referrals.csv")
postcode_population <- read_csv("postcode_population.csv")
```


```{r}
# Function to check if all values in a vector are equal
is_constant <- function(x) {
  isTRUE(all.equal(x, rep(x[1],length(x))))
}
```


```{r}
# Function to remove columns containing only missing values or all values equal
remove_meaningless_columns <- function(object) {
  all_miss <- colSums(is.na(object)) == NROW(object)
  all_equal <- apply(object, 2, is_constant)
  return(object[,!all_miss & !all_equal])
}
```


```{r}
# Create main tables as tibbles
referrals_clean1 <- referrals %>%
  janitor::clean_names() %>%
  remove_meaningless_columns() %>%
  arrange(date_received)%>%
  mutate(client_type = str_to_lower(client_type))
```
Then I checked the data type, name and value in each variable:
```{r}
#check variables' type
glimpse(referrals_clean1)
```
After checking that, I renamed the variable and combined the catogorical variables' values and created referrals_clean2.
```{r}
#change variable name
referrals_clean1<-referrals_clean1%>%rename(post_code=we_suggest_you_do,region=postcode)
```

```{r}
# combining
referrals_clean2<-referrals_clean1%>%
  mutate(client_type = recode(client_type, beareaved = "bereaved person", witnes= "witness","witnes on scene"="witness", .missing = "unknown"))%>%
  mutate(referred_by = recode(referred_by, .missing = "unknown"))
  
referrals_clean2 <- referrals_clean2 %>%
  mutate(
    client_type = if_else(str_detect(client_type, "bereaved"), "bereaved", client_type)
  ) %>%
  mutate(
    client_type = if_else(str_detect(client_type, "parent|fam|fr|mother|partner|work|employ"), "fam/fr of casualty", client_type)
  ) %>%
  mutate(
    client_type = if_else(str_detect(client_type, "driver"), "driver", client_type)
  ) %>%
  mutate(
    client_type = if_else(str_detect(client_type, "scooter|cyc|motor"), "rider", client_type)
  ) %>%
  mutate(
    client_type = if_else(str_detect(client_type, "injure|rage"), "other injured person", client_type)
  ) %>%
  mutate(
    client_type = if_else(str_detect(client_type, "witness|scene"), "witness", client_type)
  ) %>%
  mutate(
    client_type = if_else(str_detect(client_type, "pass|pedes"), "passenger", client_type)
  ) %>%
  mutate(
    client_type = if_else(str_detect(client_type, "ccccc|group|emerg|second|inap"), "other", client_type)
  ) %>%
  mutate(
    referred_by = if_else(str_detect(referred_by, "General|Hospital|Social|Internet|GP|Ambulance|RTSSV|Solicitor|After|Agencies|Counsellor|Media|Metro"), "Other", referred_by)
  )
```

Next I created a new variable "day_of_week" to check whether more referrals are received on Monday

```{r}
referrals_clean3<-referrals_clean2%>%
  mutate(referred_by=str_to_lower(referred_by))%>%
  select(-x1)%>%
  mutate(day_of_week = wday(date_received, label = TRUE)) #, post_code = as.numeric(post_code)  can not find the place 8344 gipsland
dowtb<-referrals_clean3%>%group_by(day_of_week)%>%summarise(proportion=n()/8037*100)
dowtb%>%kable()
```

\@ref(fig:dow-plot)

```{r dow-plot, fig.cap="proportion of days of week", echo=FALSE}
p_dow<-dowtb%>%
  ggplot(aes(x=day_of_week, y=proportion))+
  geom_col()
ggplotly(p_dow)
```
According to the figure \@ref(fig:dow-plot) , there are more referrals on Monday than on other days of week. This might be because that usually some referrals on weekend are postponed to the following day, which is Monday, and this can also explain why Tuesday has second largest amount of referrals, and the rest of the week days(Wednesday, Thursday and Friday) have about the same amount, which is still significantly larger(around 40%) than weekend.

We can also have a look at the summary of client type variable

```{r client, echo=FALSE}
# check the results  
client_type_count<-referrals_clean2%>%
  count(client_type)
referredby_count<-referrals_clean2%>%
  count(referred_by)  
client_type_count%>%arrange(desc(n))%>%kable(caption="the summary of client type")
```
From the table \@ref(tab:client) above we can see that the major source of client is from driver, witness and bereaved people.


and referred_by variable
```{r referredby-tb}
referredby_count%>%arrange(desc(n))%>%kable()
```

And it is clear that VPeR is the major source of referrals.

#### time related analysis

```{r}
ts_month_week<-referrals_clean3%>%
mutate(date_received = if_else(is.na(date_received), date_entered, date_received))%>%
  filter(!is.na(date_received))%>%
  mutate(month = yearmonth(date_received))%>%
  filter(month != yearmonth("2016 Jun") & month != yearmonth("2022 Apr"))%>%group_by(month, day_of_week)%>%
  summarise(referrals = n())
ts_month_week%>%
  ggplot(aes(x = month, y = referrals, color = day_of_week)) +
  geom_line()
```
Firstly as mentioned before, \@ref(fig:dow-plot) indicates that there may be a seasonal pattern in weekly data, and this plot above shows whether the pattern change over time and we can see that the pattern seems to be almost unchanged all the time. 

```{r weekly-nreferrals, fig.cap="number of referrals per week"}
ts_week<-referrals_clean3%>%
  filter(!is.na(date_received))%>%
  mutate(week = yearweek(date_received))%>%
  group_by(week)%>%summarise(referrals = n())%>%
  filter(week != yearweek("2016 Jun") & week != yearweek("2022 Apr"))%>%
  as_tsibble(index = week)
ts_week%>%fill_gaps()%>%
  autoplot()+
  geom_smooth(span = 0.4)

```
we can see a cyclic bahaviour from the start till 2020, and a significant drop in 2020 and then go back slowly to the cyclic pattern.


```{r weekly-season}
ts_week%>%fill_gaps()%>%
  gg_season()
```

Although a bit messy, we can see from the plot /@ref(fig:weekly-nreferrals) that the number of referrals is usually lower in winter than that in summer.
It will be clearer in monthly plot /@ref(fig:monthly-season):

```{r monthly-season,fig.cap="number of referrals per month"}
ts_month<-referrals_clean3%>%
mutate(date_received = if_else(is.na(date_received), date_entered, date_received))%>%
  filter(!is.na(date_received))%>%
  mutate(month = yearmonth(date_received))%>%
  group_by(month)%>%summarise(referrals = n())%>%
  filter(month != yearmonth("2016 Jun") & month != yearmonth("2022 Apr"))%>%
  as_tsibble(index = month)
ts_month%>%fill_gaps()%>%
  gg_season()
```
Maybe the drop in the 2nd quarter is because the staff in the organization usually have a vocation due to the cold weather in the 2nd quarter.

```{r sub-month, fig.cap="subseries of monthly data"}

ts_month%>%fill_gaps()%>%
  gg_subseries()

```
From the plot /@ref(fig:sub-month) above we can see that almost every month there is a significant drop in the year 2020, which can be the consequence of Australia's lock-down policy for COVID-19.

This graph shows subseries of the referrals data by month. As is shown above the peaks in terms of the number of referrals are in March and October, and the troughs are in May and Sep. Most months show a downward trend from 2017 to 2022 but have troughs in 2020 or 2021, which may be a consequence of covid-19. But overall the seasonality is weak as the ETC function suggests an MNN model which does not contain a seasonal component.


```{r}
ts_month_1<-referrals_clean3%>%
mutate(date_received = if_else(is.na(date_received), date_entered, date_received))%>%
  filter(!is.na(date_received))%>%
  mutate(month = yearmonth(date_received))%>%
  group_by(month,client_type)%>%summarise(referrals = n())%>%
  filter(month != yearmonth("2016 Jun") & month != yearmonth("2022 Apr"))
p<-ts_month_1%>%
  ggplot( aes(month, referrals,fill=client_type), position="fill") +
  geom_col()

ggplotly(p)
```


```{r}
ts_month_referredby<-referrals_clean3%>%
mutate(date_received = if_else(is.na(date_received), date_entered, date_received))%>%
  filter(!is.na(date_received))%>%
  mutate(month = yearmonth(date_received))%>%
  group_by(month,referred_by)%>%summarise(referrals = n())%>%
  filter(month != yearmonth("2016 Jun") & month != yearmonth("2022 Apr"))
p1<-ts_month_referredby%>%
  ggplot( aes(month, referrals,fill=referred_by), position="fill") +
  geom_col()

ggplotly(p1)
```

As is shown above, the referrals from vsa disappeared since 2019 Sep.

```{r}
fit<-ts_month%>%fill_gaps()%>%model(etsmonth = ETS(referrals))
report(fit)
```

```{r}
referrals_clean3%>%
  filter(!is.na(date_received)&!is.na(date_entered))%>%
  mutate(gap = date_entered-date_received)%>%
  filter(gap>-1&gap<8&date_received>"2020-03-01")%>%
  group_by(date_received)%>%
  summarise(gap=mean(gap))%>%
  as_tsibble(index = date_received)%>%
  autoplot(gap)+
  geom_smooth(span=.4)
```

Although not very apparent, we can still observe the downward pattern followed by an upwards pattern in this gap plot, combining with the drop of referrals after the outbreak of covid-19, we can infer that the number of referrals does to some degree influence the speed of processing referrals, and this indicates that the workload might be close to the upper limit of RTSSV.



### spacial data analysis

```{r}
postcode_population<-postcode_population%>%mutate(postcode=as.character(postcode))
```
```{r}
referrals_clean4<-referrals_clean3%>%
  group_by(post_code)%>%
  summarise(referrals=n())
mapdata<-postcode2021%>%
  left_join(referrals_clean4, by=c("postcode_2021"="post_code"))%>%
  left_join(postcode_population, by=c("postcode_2021"="postcode"))%>%
  select(cent_lat, cent_long, population, referrals, postcode_2021,geometry)
  
```



```{r}
mapdata[is.na(mapdata)] <- 0
mapdata1<-mapdata%>%
  summarise(postcode=postcode_2021,attractiveness=referrals/sum(referrals)-population/sum(population), lat = cent_lat, lng = cent_long, geometry=geometry,population=population, referrals=referrals)%>%mutate(attractiveness=1000*attractiveness, postcode=as.numeric(postcode))%>%filter(postcode>2999&postcode<4000)
```





```{r map}
map <-
mapdata1 %>%
  ggplot() +
  geom_sf(aes(geometry = geometry, fill=attractiveness, text=paste("postcode:",postcode, "\n", "population:", population, "\n", "referral:",referrals)),lwd=0, show.legend = TRUE)+  # use the geometry variable
  geom_point(aes(145.2,-37.8),color = "red",size=0.6)+
  scale_fill_viridis_c() +

  coord_sf()

ggplotly(map)
```





To make comparasion between the population distribution and the distrbution of the source of referrals, I created a variable calls attractiveness, which is computed by the formula below:
$$\frac{a_1}{a_2}-\frac{b_1}{b_2}$$
In this formula, $a_1$ stands for the number of referrals in a certain postcode zone, $a_2$ stands for the number of referrals in the whole area, $b_1$ stands for the population in a certain postcode zone and $b_2$ stands for the population in the whole area.

The reason I didn’t use a1/b1 is the problem with robustness when there is zero referral because in some areas the number of referrals(a1) is zero, therefore a1/b1 is also zero, but the same “zero” can have different meaning since the size of the population in that certain area(b1) can vary a lot.


```{r}
ts_month_1<-referrals_clean3%>%
mutate(date_received = if_else(is.na(date_received), date_entered, date_received))%>%
  filter(!is.na(date_received))%>%
  mutate(month = yearmonth(date_received))%>%
  group_by(month,client_type)%>%summarise(referrals = n())%>%
  filter(month != yearmonth("2016 Jun") & month != yearmonth("2022 Apr"))
p<-ts_month_1%>%
  ggplot( aes(month, referrals,fill=client_type), position="fill") +
  geom_col()

ggplotly(p)
```












